---
title: "JPM Predictive Model"
author: "Thao Nguyen Pham"
date: "4/18/2023"
output: html_document
---
#### **1. PREPARE THE DATASET**

1.1. Import the dataset

* S&P500 dataset 

```{r}
sp500 <- read.csv("C:/Users/PC/Desktop/Project/PredictiveModeling/sp500-1022.csv")
head(sp500)
```

* JPMorgan Chase dataset 

```{r}
jpm <- read.csv("C:/Users/PC/Desktop/Project/PredictiveModeling/jpm-1022.csv")
head(jpm)
```

* US 3-month treasury bills dataset 

```{r}
US3M <- read.csv("C:/Users/PC/Desktop/Project/PredictiveModeling/US3M.csv", fileEncoding = 'UTF-8-BOM')
head(US3M)
```

* Economic indicators dataset 

```{r}
indicator <- read.csv("C:/Users/PC/Desktop/Project/PredictiveModeling/indicators.csv", fileEncoding = 'UTF-8-BOM')
head(indicator)
```

1.2. CALCULATE MARKET QUARTERLY RATE OF RETURN 

```{r}
#create function to calculate quarterly simple return rate
quarterly_return <- function (df) {
  
  df_quarterly_return <- df #create a copy of the input dataframe
  
  #loop through dataframe to calculate quarterly return, in which current quarter return = ((current price - previous price)/previous price)*100
  for (c in 2:ncol(df_quarterly_return)) {
    for (r in 2:nrow(df_quarterly_return)) {
      df_quarterly_return[r, c+1] <- as.numeric(format((df[r, c] - df[r-1, c])/df[r-1, c]), digits = 4)
    }
    df_quarterly_return[1, c+1] = 0 # set the value of the first row = 0 
  }
  return (df_quarterly_return)
}
```

* Calculate market quarterly simple rate of return
```{r}
sp500_qr <- quarterly_return(sp500)[-1, ]
colnames(sp500_qr) <- c("Date", "sp500_price", "sp500_quarterlyReturn")
head(sp500_qr)
```

* Calculate JPM stock's quarterly simple rate of return
```{r}
jpm_qr <- quarterly_return(jpm)[-1, ]
colnames(jpm_qr) <- c("Date", "jpm_price", "jpm_quarterlyReturn")
head(jpm_qr)
```

1.3. Merge the quarterly return datasets into 1 table

```{r}
#Merge sp500_qr & jpm_qr & US3M together
ds <- merge(sp500_qr, jpm_qr, by = c("Date" = "Date"), all.x = TRUE)
ds
```

1.4. Find Beta

** Reference for computing the betas here: https://www.wikihow.com/Calculate-Beta#:~:text=Find%20the%20coefficient%20for%20the,x%20value%20is%20your%20beta.**

```{r}
ds$beta <- c(ds$sp500_quarterlyReturn - US3M$Risk.free.Rate)/(ds$jpm_quarterlyReturn - US3M$Risk.free.Rate)
ds
```

#### **II. Development of statistical model**

2.1. Check the normality of the Quarterly Closing Price (QCP)

```{r}
#draw Q-Q plot to check the normality of QCP
qqnorm(ds$jpm_price, pch = 1, frame = FALSE)
qqline(ds$jpm_price, col = "steelblue", lwd = 2)
```

```{r}
#perform Shapiro-Wilk test for normality
shapiro.test(ds$jpm_price) 
#p-value < 0.05 => not normal distributed
```

2.2. AFTER JOHNSON TRANSFORMATION ON MINITAB STATISTICAL SOFTWARE

Enter the parameter values from Johnson Transformation
γ = 0.636034, δ = 0.613498, 𝜉 = 20.8924 and 𝜆 = 144.6576

```{r}
#Import transformed JPM QCP
jpm_trans <- read.csv("C:/Users/PC/Desktop/Project/PredictiveModeling/jpm_trans.csv", fileEncoding = 'UTF-8-BOM')
colnames(jpm_trans)[2] <- "QCP_T"
jpm_trans
```

```{r}
#draw Q-Q plot to check normality of transformed JPM QCP
qqnorm(jpm_trans$QCP_T, pch = 1, frame = FALSE)
qqline(jpm_trans$QCP_T, col = "steelblue", lwd = 2)
```

```{r}
#perform Shapiro-Wilk test for normality after Johnson transformation
shapiro.test(jpm_trans$QCP_T) 
```

2.3. CHECK CORRELATION USING DATA TRANSFORMATION

```{r}
library(corrplot)
```

```{r}
#Generate the dataset that contains all indicators and transformed JPM QCP
df <- Reduce(function(...) merge(..., all = TRUE, by = "Date"),
       list(ds[, c(1, 6)], indicator, jpm_trans))
df
```

```{r}
#Check the correlation among the indicators
corrplot(cor(df[, c(2:12)]), type = "upper", addCoef.col="orange", number.cex=0.75)
```

2.4. VARIANCE INFLATION FACTOR

```{r}
library(car)  
```

```{r}
# Check Variance Inflation Factor (VIF) to verify the existence of multicollinearity among attributable indicators
#fit the regression model
model <- lm(QCP_T ~ beta + PB + PE + FCF.Share + PEG + 
            Dividend.Yield + GDP + Interest.Rate + CPI + PSR,
            data = df)

#view the output of the regression model
summary(model)

#calculate the VIF for each predictor variable in the model
vif(model)
```

DROP HIGH VIF SCORES FACTORS > 10 (Which is GDP & PB)
```{r}
#fit the regression model
model <- lm(QCP_T ~ beta + PE + FCF.Share + PEG + 
            Dividend.Yield + Interest.Rate + CPI + PSR,
            data = df)

#view the output of the regression model
summary(model)

#calculate the VIF for each predictor variable in the model
vif(model)
```

2.5. LASSO REGRESSION

Prepare data for lasso regression
```{r}
library(caTools)
library(caret)
library(dplyr)
```

```{r}
#create a dataset containing only indicators for training and testing purpose
y <- as.matrix(df$QCP_T)    #response variable
x <- model.matrix(QCP_T ~ .^2, df[, -1])[, -1]   #predictor variables


data <- data.frame(x, QCP_T = y)

# Set the proportion of data to use for training
set.seed(1)
train_prop <- 0.8

# Split the data into 80% training and 20% testing sets
train_rows <- sample(1:nrow(data), round(train_prop * nrow(data)), replace = FALSE)
train_data <- data[train_rows, ]
test_data <- data[-train_rows, ]


y_train <- as.matrix(train_data$QCP_T)
x_train <- as.matrix(select(train_data, -QCP_T))

y_test <- as.matrix(test_data$QCP_T)
x_test <- as.matrix(select(test_data, -QCP_T))

```

Find optimal lambda value for lasso regression
```{r}
library(glmnet)
```

```{r}
# Perform k-fold cross-validation to find optimal lambda value (default nfolds = 10)
lasso_reg <- cv.glmnet(x_train, y_train, alpha = 1,standardize = TRUE) 
plot(lasso_reg)

#find optimal lambda value that minimizes test MSE
best_lambda <- lasso_reg$lambda.min
best_lambda
```

Train LASSO model
```{r}
# Fit the lasso model with the optimal lambda value
lasso_model <- glmnet(x_train, y_train, alpha = 1, lambda = best_lambda, standardize = TRUE)

# Compute R^2 from fact and predicted values
eval_results <- function(fact, predicted, df) {
  SSE <- sum((predicted - fact)^2)
  SST <- sum((fact - mean(fact))^2)
  R_square <- 1 - SSE / SST
  RMSE <- sqrt(SSE/nrow(df)) 
  
  print(paste("SSE:", SSE))
  print(paste("SST:", SST))
  print(paste("R_square:", R_square))
  print(paste("RMSE:", RMSE))
  print("--------------")
  }

#Predict and evaluate lasso model using training dataset
predict_train <- predict(lasso_model, s = best_lambda, newx = x_train)
eval_results(y_train, predict_train, y_train)

#Predict and evaluate lasso model using testing dataset
predict_test <- predict(lasso_model, s = best_lambda, newx = x_test)
eval_results(y_test, predict_test, y_test)

#Fit lasso model on full dataset
out <- glmnet(x, y, alpha = 1) # Display coefficients using lambda chosen by CV
lasso_coef <- coef(out, s = best_lambda)
plot(out, xvar = "lambda", label = TRUE)
print (lasso_coef)
```

2.6. BACKWARD ELIMINATION
```{r}
library(MASS)
```

FIT MODEL WITH IMPORTANT INDICATORS & INTERACTIONS FROM LASSO 
```{r}
important_after_lasso <- c("PB", "PE", "PEG", "CPI", "FCF.Share", "Dividend.Yield", "GDP", "Interest.Rate" ,
               "PE.Interest.Rate", "FCF.Share.PEG", "FCF.Share.Interest.Rate", "FCF.Share.CPI", 
               "PEG.Dividend.Yield", "PEG.Interest.Rate")
train_data_1 <- train_data[, important_after_lasso]

#fit model based on important indicators and interactions from LASSO
model <- lm(train_data$QCP_T ~ ., data = train_data_1)
summary(model)
```

APPLY BACKWARD ELIMINATION

```{r}
#Apply backward elimination
step.model <- stepAIC(model, direction = "backward", trace = TRUE)
summary(step.model)
a <- summary(step.model)
```

3. MODEL EVALUATION

```{r}
#predict the model
predict_model_train <- predict(step.model, newdata = data.frame(x_train))
predict_model_test <- predict(step.model, newdata = data.frame(x_test))

#calculate mean & sum of residuals
mean(a$residuals)
sum(a$residuals)

#Calculate the RMSE
rmse_train <- sqrt(mean((y_train - predict_model_train )^2))
rmse_train

rmse_test <- sqrt(mean((y_test - predict_model_test)^2))
rmse_test

#Calculate the RRMSE
rrmse_train <- rmse_train/diff(range(y_train)) * 100
rrmse_train

rrmse_test <- rmse_test/diff(range(y_test)) * 100
rrmse_test

#Calculate the MAPE
mape_train <- mean(abs((y_train - predict_model_train) / y_train)) * 100
mape_train

mape_test <- mean(abs((y_test - predict_model_test) / y_test)) * 100
mape_test

#Test for normality of residuals
shapiro.test(step.model$residuals)
```

```{R}
#Test for autocorrelation
durbinWatsonTest(step.model)
plot(step.model)
```